---
title: "BIS15L_AI_class"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?
AI, or Artificial Intelligence, refers to the simulation of human intelligence processes by machines, especially computer systems. These processes include learning (the acquisition of information and rules for using the information), reasoning (using rules to reach approximate or definite conclusions), and self-correction. AI encompasses a variety of approaches and techniques, including machine learning, natural language processing, computer vision, robotics, and more. The ultimate goal of AI is to create systems that can perform tasks that would typically require human intelligence, such as understanding natural language, recognizing patterns, making decisions, and solving problems.


2. What is ChatGPT?
ChatGPT is an AI-based conversational agent developed by OpenAI. It is built on the GPT (Generative Pre-trained Transformer) architecture, specifically GPT-3.5 in my case, which is a state-of-the-art language model capable of understanding and generating human-like text based on the input it receives. ChatGPT is designed to engage in natural language conversations, answer questions, provide information, generate text, and perform various other language-related tasks. It can adapt to a wide range of topics and contexts, making it useful for a variety of applications, including customer service, language translation, education, and more.

3. What is Github Copilot?
GitHub Copilot is a tool that helps programmers write code faster by giving them suggestions as they type. It uses artificial intelligence to understand what code you're working on and offers helpful completions and snippets to speed up the process. It's like having a smart assistant for coding tasks.

4. How to use Rstudios?
Sure thing! Here's a more casual summary:

Get R and RStudio installed on your computer.
Open up RStudio and check out the different sections like the Console and the File Explorer.
Start typing your R code in the Console or in a new script file.
If you need extra tools or functions, you can install and load them with a couple of commands.
If you're stuck or need help, RStudio has a Help section to guide you.
And remember to save your work regularly, just like you would with any other program!
That's the gist of it! Now you're ready to dive into some R coding fun.

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix

netflix <- read_csv("data/netflix_titles.csv")
```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
str(netflix)
head(netflix)
summary(netflix)
glimpse(netflix)
```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*
#I just asked it nicely to do it how I like it 

Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
netflix_movies <- filter(netflix, type == "Movie")
head(netflix_movies)
```


Which director has the most movies on netflix?
```{r}
director_counts <- netflix_movies %>%
                  group_by(director) %>%
                  count() %>%
                  arrange(desc(n))
head(director_counts)
```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

Filter the directors with the most movies?
```{r}
netflix_movies_filtered <- netflix_movies %>%
                           filter(!is.na(director))

director_counts <- netflix_movies_filtered %>%
                   group_by(director) %>%
                   count() %>%
                   arrange(desc(n))
```

```{r}
top_10_director <- head(director_counts, 10)
```

*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


Separate each director into their own row then show the directors with the most movies
```{r}
netflix_movies <- netflix_movies %>%
                separate_rows(director, sep = ", ")

count(netflix_movies, director, release_year, sort=TRUE) %>%  filter(!is.na(director))
```


Make a bar plot of the top directors with the most movies on netflix
```{r}
ui <- fluidPage(
  titlePanel("Top 10 Directors with the Most Movies on Netflix"),
  mainPanel(
    plotOutput("barplot")
  )
)

# Define server logic
server <- function(input, output) {
  
  # Render the bar plot
  output$barplot <- renderPlot({
    # Filter top 10 directors
    top_10_director <- head(top_10_director, 10)
    
    # Create bar plot
    ggplot(top_10_director, aes(x = reorder(director, num_movies), y = num_movies)) +
      geom_bar(stat = "identity", fill = "skyblue") +
      labs(title = "Top 10 Directors with the Most Movies on Netflix",
           x = "Director",
           y = "Number of Movies") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

[Insert your prompt here]
```{r}

```

*How about for just 2020*



[Insert your prompt here]
```{r}

```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt
```{r}
ui <- fluidPage(
  titlePanel("Top 10 Directors with the Most Movies on Netflix"),
  sidebarLayout(
    sidebarPanel(
      # No sidebar input needed
    ),
    mainPanel(
      # Output: Plot the top directors bar chart
      plotOutput("top_directors_plot")
    )
  )
)

# Define server logic
server <- function(input, output) {
  
  # Render the top directors bar chart
  output$top_directors_plot <- renderPlot({
    # Filter the top 10 directors
    top_10_directors <- top_director %>%
                        slice_max(order_by = num_movies, n = 10)
    
    # Create the bar plot
    ggplot(top_10_directors, aes(x = reorder(director, num_movies), y = num_movies)) +
      geom_bar(stat = "identity", fill = "skyblue") +
      labs(title = "Top 10 Directors with the Most Movies on Netflix",
           x = "Director",
           y = "Number of Movies") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      coord_flip()
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```


*Now use ChatGPT, Copilot or previous notes to fix the code*





## That's it! Let's take a break and then move on to working on group projects.
