---
title: "lab5_warmup"
author: "Pachia Lee"
date: "2024-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## Warm-up
1. Load the `bison.csv` data.

```{r}
bison <- read.csv("data/bison.csv")
```

2. What are the dimesions and structure of the data?

```{r}
dim(bison)
str(bison)
```


3. We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.

```{r}
bison%>%
  select(animal_code, animal_sex, animal_weight, animal_yob)
```



```{r}
new_bison<- select(bison,"animal_code", "animal_sex", "animal_weight", "animal_yob")
new_bison
```

4. Pull out the animals born between 1980-1990.

```{r}
new_bison<- select(bison,"animal_code", "animal_sex", "animal_weight", "animal_yob")%>%
  filter(animal_yob>= "1980" & animal_yob<= "1990")
new_bison
```

5. How many male and female bison are represented between 1980-1990?

```{r}
males <- new_bison%>%
  filter(animal_yob>=1980 & animal_yob<= 1990)%>%
  filter(animal_sex=="M")
males

table(males$animal_sex)
```

```{r}
females <- new_bison%>%
  filter(animal_yob>=1980 & animal_yob<= 1990)%>%
  filter(animal_sex=="F")
females

table(females$animal_sex)
```

6. Between 1980-1990, were males or females larger on average? #females are larger on average

```{r}
mean(males$animal_weight)
```

```{r}
mean(females$animal_weight)
```


